### 2008-07-01 trunk/Papers/rmrs-08/introduction.tex
restructured intro

% What we'll do here.  Problem: RMRS is designed to support building
% semantic components to shallow parsers, and is intended to offer a way
% of integrating semantic representations
% computed by deep and shallow parsing.  We demonstrate the feasibility
% of this for the first time, by introducing a model theory for RMRS,
% defining entailment among RMRS representations and characterising this
% entailment relation 
% syntactically as extension of solved forms (ref to solved forms).  
% We show that the ``compatability'' of
% deep and shallow semantic representations is definable in terms of the
% entailment 
% relation between the RMRS
% representations.

% just so we have a citation here: \cite{Copestake_etal:MRS}

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.
These need access to the vast array of
semantic information that is captured in conventional logical forms,
including (but not limited to) information about semantic scope and
predicate argument structure.  Hand-crafted grammars yield detailed
logical forms (e.g., \cite{butt:etal:1999,copestake:flickinger:2000}), but deep parsing tends to lack
robustness: hand-crafted grammars fail to cover all words and
linguistic constructions; and unedited text and speech contains
ill-formed phrases, which by design deep grammars do not handle.  
\hidden{
Deep
parsers on their own also lack ambiguity resolution techniques,
producing all possible analyses.  Integration of statistical parse
selection with deep parsing is promising (e.g.,
\cite{toutanova:etal:2004}), but is not a solved problem.
}

@@

This paper focusses on a particular approach to producing semantic
information from robust parsers, exemplified in
\cite{copestake:2003,frank:2004}, among others.  Their strategy is to
utilise semantic underspecification to semi-automatically build
semantic components to shallow parsers, so that the output neither
over-determines nor under-determines the semantic information that is
revealed by the (shallow) syntactic analysis.  The semantic formalism
used to express this is Robust Minimal Recursion Semantics ({\sc
  rmrs}, \cite{copestake:2003}); this is an extension of {\sc mrs}
\cite{copestake:etal:2005} that is designed to be maximally flexible
in the type of semantic information that can be left underspecified:
it can express partial information about semantic scope, the values of
arguments to predicate symbols and/or their argument position, the
arity of the predicate symbols and the sorts of arguments they take.
We show in Section~\ref{sec:motivation} that all these features are
needed when information about lexical subcategorisation or
syntactic dependencies is missing---a characteristic feature of
shallow parsers.  Several researchers have demonstrated that {\sc
  rmrs} is a suitable framework on which to semi-automatically
construct semantic components to shallow parsers, ranging in depth
from {\sc pos} taggers (refs) to chunk parsers (frank ref) and
intermediate parsers ({\sc rasp} ref).

>>>>

Representing semantics as a logical form that supports automated
inference and model generation is vital for deeper language
engineering tasks, such as dialogue systems.  Such detailed semantic
representations can be obtained from hand-crafted deep grammars (e.g.,
\cite{butt:etal:1999,copestake:flickinger:2000}), but deep parsing
tends to lack robustness: hand-crafted grammars fail to cover all
words and linguistic constructions; and unedited text and speech
contains ill-formed phrases, which by design deep grammars do not
handle.

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\cite{bos:etal:2004} build upon a statistical CCG parser, exploiting
the elegant syntax-semantics interface that CCG provides;
\cite{zettlemoyer:collins:2007} and others combine syntactic analyses
with shallow features to learn string-to-semantics mappings; and
\todo{say something about RASP and Anette's work}.

However, there are certain semantic phenomena that such shallow
approaches can't capture, including quantifier scope, optional
arguments, and long-distance dependencies.  Forcing a shallow parser
to make a decision about these phenomena is therefore a major source
of errors.  It would be preferrable to give the parser the option to
declare itself not sufficiently informed to make the decision, and
leave the decision open -- i.e., to compute a partial semantic
representation that neither over-determines nor under-determines the
semantic information revealed by the (shallow) syntactic analysis, and
can later be completed by further inferences.

In this paper, we focus on one particular approach to representing
partial semantic information and producing it from robust parsers:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005} designed to support underspecification both
of scope relations and of the predicate-argument structure, and has
been demonstrated to support semantic constructions of shallow parsers
ranging in depth from {\sc pos} taggers (refs) to chunk parsers (frank
ref) and intermediate parsers ({\sc rasp} ref).  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
\cite{copestake:flickinger:2000} can also be seen as special cases of
{\sc rmrs} representations.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other framework such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by a
partial \rmrs.

We will first illustrate the differences and challenges with deep and
shallow semantic construction in an example in
Section~\ref{sec:motivation}.  We will then define the syntax and
model-theoretic semantics of \rmrs\ and introduce some important
formal concepts in Section~\ref{sec:rmrs}.  Finally, we demonstrate
the practical usefulness of our formalisation in
Section~\ref{sec:entailment} by modelling compatibility of
partial semantic representations as \rmrs\ entailment and providing a
syntactic criterion for characterising entailment.
Section~\ref{sec:conclusion} concludes and points to future work.

\hidden{

The language in which these partial representations are expressed

Moreover, in a hybrid system that attempts to combine the outputs of
deep and shallow parsers,

@@

This paper focuses on a particular approach to producing partial
semantic information from robust parsers, exemplified in
\cite{copestake:2003,frank:2004}, among others.  Their strategy is to
utilise semantic underspecification to semi-automatically build
semantic components to shallow parsers, so that the output neither
over-determines nor under-determines the semantic information that is
revealed by the (shallow) syntactic analysis.  The semantic formalism
used to express this is Robust Minimal Recursion Semantics

; this is a generalisation of

 that is designed to be maximally flexible
in the type of semantic information that can be left underspecified:
it can express partial information about semantic scope, the values of
arguments to predicate symbols and/or their argument position, the
arity of the predicate symbols and the sorts of arguments they take.
We show in Section~\ref{sec:motivation} that all these features are
needed when information about lexical subcategorisation or syntactic
dependencies is missing---a characteristic feature of shallow parsers.
Several researchers have demonstrated that {\sc rmrs} is a suitable
framework on which to semi-automatically construct semantic components
to shallow parsers, ranging in depth .
